\documentclass[journal]{IEEEtran}

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{array}
\usepackage{url}
\usepackage[hidelinks]{hyperref}
\usepackage{orcidlink}

\begin{document}

\title{Dual-Representation Mondrian Conformal Assignment of Pressure-Transient
GeoTypes: A Reproduction and Extension for Fractured-Reservoir Flow}

\author{%
  Felipe Santib\'a\~nez-Leal~\orcidlink{0000-0002-0150-3246},~\IEEEmembership{Independent Researcher}%
  \thanks{F.\ Santib\'a\~nez-Leal is with the CAOS open-research programme,
    Santiago, Chile (e-mail: \texttt{fsantibanezleal@ug.uchile.cl};
    ORCID: 0000-0002-0150-3246).}%
  \thanks{Preprint, 2026. Code, committed derived artifacts, and the interactive
    workbench are released under MIT at
    \url{https://github.com/fsantibanezleal/CAOS_RES_Pulso}; the reusable shape
    machinery is the Apache-2.0 package \texttt{pygeotypes}
    (\url{https://github.com/fsantibanezleal/CAOS_GeoTypes}). No GPL workflow code
    or raw licensed data is redistributed.}%
}

\markboth{Preprint, 2026 --- CAOS open-research programme}{Santib\'a\~nez-Leal:
Dual-Representation Mondrian Conformal Assignment of Pressure-Transient GeoTypes}

\IEEEtitleabstractindextext{%
\begin{abstract}
Fractured reservoirs exhibit recurring fluid-flow behaviours hidden in the shape of
their pressure-transient responses. A recent study clusters geologically consistent
discrete-fracture-network (DFN) ensembles into flow-behaviour classes and attributes
each class to fracture-network properties. We first reproduce that pipeline, DFN
pressure transients, Bourdet-derivative features, dynamic-time-warping (DTW)
$k$-medoids clustering into a GeoType catalogue, and random-forest / TreeSHAP
attribution, on the original open 4TU corpus and on analytic dual-porosity families,
recovering clean clusters (silhouette up to $0.86$ on the real transients, higher than
on the analytic families) and the finding that fracture intensity ($\log I$) is the
top attributed control. We then extend it in two ways. First, we make the assignment
\emph{coverage-controlled}, which the base random-forest attribution is not, with a
class-conditional (Mondrian) split-conformal assigner on a DTW nonconformity score
that emits per-class $p$-values, a prediction set, and an out-of-catalogue flag
(empirical coverage $0.85$--$0.87$ at target, out-of-catalogue rate $0.03$--$0.15$,
mean set size $1.5$--$1.7$). Second, and the methodological contribution, we introduce
a \emph{dual-representation Mondrian conformal} assigner that conformalizes jointly
across two nonconformity scores, the DTW shape distance and a physics-descriptor
implausibility from the attribution forest, and accepts a GeoType only in the
per-class conjunction of the two prediction sets. This excludes curves that are
shape-close to the wrong physics, which single-score conformal accepts: on the
rich-method cases the dual conjunction tightens the mean set size below one and flags
right-shape-wrong-physics curves at an honest cost of coverage (e.g.\ $0.85\to0.73$,
$0.87\to0.75$). The shape machinery is signal-agnostic and is exercised on
hydrogeology pumping-test field data as well as reservoir transients. The
contribution is a faithful reproduction plus a coverage-controlled assignment that
fuses shape space and physics-descriptor space, with every number backed by a
committed deterministic artifact.
\end{abstract}

\begin{IEEEkeywords}
fractured reservoirs, pressure-transient analysis, discrete fracture networks,
dynamic time warping, k-medoids, conformal prediction, Mondrian, random forests,
SHAP, reproducibility, geothermal, hydrogeology.
\end{IEEEkeywords}}

\maketitle
\IEEEdisplaynotcompsoctitleabstractindextext
\IEEEpeerreviewmaketitle

% =====================================================================
\section{Introduction}
% =====================================================================

\IEEEPARstart{T}{he} pressure-transient response of a well is a fingerprint of the
reservoir's flow structure: its shape, and the shape of its Bourdet
derivative~\cite{bourdet1989}, distinguishes homogeneous from dual-porosity behaviour
and localizes the interporosity transition~\cite{warren1963}. In fractured reservoirs
the mapping from geology to transient is many-to-one and uncertain, so a natural
question is whether the space of transients organizes into a small catalogue of
recurring flow behaviours, a ``geological DNA'', and if so, which fracture-network
properties control membership. A recent study answers both by clustering geologically
consistent discrete-fracture-network (DFN) ensembles into flow-behaviour classes and
attributing them~\cite{kameltarghi2026}.

This paper reproduces that pipeline and extends it with a coverage-controlled
assignment. Reproduction matters here for two reasons: the original workflow depends
on a GPL-licensed code base we deliberately do not vendor, and a clustering-plus-attribution
result is only trustworthy if an independent re-implementation on the same open data
recovers it. We re-implement the shape machinery from permissively licensed components
(DTW~\cite{sakoe1978}, $k$-medoids, silhouette selection~\cite{rousseeuw1987}, random
forests~\cite{breiman2001}, TreeSHAP~\cite{lundberg2020}) as the open package
\texttt{pygeotypes}, run it on the original 4TU corpus~\cite{data4tu} and on analytic
dual-porosity families, and confirm the clusters and the top attributed control.

The extension addresses a gap in how such catalogues are \emph{used}. Random-forest
attribution ranks controls but gives no guarantee on a new curve's class, and a curve
can be shape-close to the wrong physics. We therefore add conformal
prediction~\cite{vovk2005,angelopoulos2023}: first a class-conditional (Mondrian)
split-conformal assigner on the shape distance, then a dual-representation variant that
conformalizes jointly over shape and physics-descriptor scores. Our contributions are:
(i) an independent, permissively licensed reproduction of the DFN-transient GeoType
pipeline that recovers its clusters and fracture-intensity attribution; (ii) a
coverage-controlled Mondrian conformal GeoType assigner; (iii) a dual-representation
Mondrian conformal assigner that excludes right-shape-wrong-physics curves, with the
coverage/tightness trade-off reported honestly.

% =====================================================================
\section{Reproduced pipeline}
\label{sec:pipeline}
% =====================================================================

A GeoType catalogue is built in four stages. \emph{Transients and features.} Each DFN
ensemble (or analytic study) yields pressure transients; from each we compute the
log-time pressure and its Bourdet derivative~\cite{bourdet1989}, and for analytic
dual-porosity references the Warren--Root response~\cite{warren1963} is evaluated by
Stehfest numerical Laplace inversion~\cite{stehfest1970}. \emph{Clustering.} Curves are
compared by dynamic time warping~\cite{sakoe1978} and grouped by $k$-medoids; the
catalogue size $K$ is chosen by a silhouette sweep~\cite{rousseeuw1987}, and each
GeoType is represented by its medoid curve and a membership. \emph{Attribution.} A
random forest~\cite{breiman2001} predicts the GeoType from physical descriptors of the
network, accuracy-gated (attribution is withheld when the forest cannot predict the
labels), and TreeSHAP~\cite{lundberg2020} ranks the controlling descriptors.
\emph{Assignment.} A new curve is classified against the baked medoids.

On the original open corpus the reproduction recovers the reported structure: the real
4TU transients cluster more cleanly than the analytic families (silhouette up to
$0.86$ versus $\approx 0.5$--$0.6$ analytic), and fracture intensity ($\log I$) is the
top attributed control (largest mean absolute SHAP), partially reproducing the paper's
fracture-intensity finding. On the analytic families the interporosity flow coefficient
$\log_{10}\lambda$ dominates the assignment sensitivity, exactly as the physics
predicts, since $\lambda$ sets the transition timing that separates the dual-porosity
behaviours.

% =====================================================================
\section{Coverage-controlled assignment}
\label{sec:conformal}
% =====================================================================

\subsection{Mondrian split-conformal on the shape score}

Random-forest attribution ranks controls but offers no guarantee that a new curve is
assigned to the right GeoType. We add a class-conditional (Mondrian) split-conformal
assigner~\cite{vovk2005,angelopoulos2023}. The nonconformity score of a curve $x$ for
GeoType $g$ is its DTW distance to the medoid, $s_\text{shape}(x,g)$. Calibrated
per class on a disjoint slice, it yields a $p$-value $p_\text{shape}(x,g)$ and a
prediction set $\{g : p_\text{shape}(x,g) > \alpha\}$; a curve matching no class is
flagged out-of-catalogue. Across the rich-method cases the empirical marginal coverage
is $0.85$--$0.87$ at the target level, the out-of-catalogue rate is $0.03$--$0.15$, and
the mean set size is $1.5$--$1.7$: a coverage guarantee the base attribution cannot give.

\subsection{Dual-representation Mondrian conformal (contribution)}

A single score cannot express that a curve may be shape-close to the wrong physics: a
transient that matches a medoid by DTW shape yet whose physical descriptors are
atypical for that GeoType. We conformalize jointly across two representations. In
addition to the shape score, define a descriptor score
\begin{equation}
s_\text{desc}(x,g) = 1 - \mathrm{RF}\big(g \mid \text{descriptors}(x)\big),
\end{equation}
the implausibility of $x$'s physical descriptors under GeoType $g$ from the attribution
forest. Both are Mondrian-calibrated, giving $p_\text{shape}$ and $p_\text{desc}$, and
the dual prediction set is the per-class conjunction
\begin{equation}
\Gamma^\alpha(x) = \big\{\, g : p_\text{shape}(x,g) > \alpha \ \wedge\
p_\text{desc}(x,g) > \alpha \,\big\}.
\label{eq:dual}
\end{equation}
A GeoType is accepted only if the curve is both shape-consistent and
descriptor-consistent; a shape match with implausible physics is excluded. This is
beyond single-score conformal and beyond random-forest/SHAP attribution, which has no
coverage control: a coverage-controlled assignment that fuses shape space and
physics-descriptor space.

Table~\ref{tab:dual} reports the honest trade-off on two rich-method cases. The dual
conjunction tightens the mean set size below one and flags right-shape-wrong-physics
curves that shape-only conformal accepts, at the cost of some marginal coverage
($0.85\to0.73$, $0.87\to0.75$). When descriptors are absent or degenerate (some real
cases), the dual layer degrades gracefully to shape-only and reports that it did.

\begin{table}[!t]
\caption{Dual-representation vs.\ shape-only Mondrian conformal. The dual conjunction
trades marginal coverage for tighter, physics-consistent sets and excludes
right-shape-wrong-physics curves.}
\label{tab:dual}
\centering
\begin{tabular}{lrr}
\toprule
Case & shape-only coverage & dual coverage (mean set) \\
\midrule
Rich-method A & 0.85 & 0.73 \; (0.78) \\
Rich-method B & 0.87 & 0.75 \; (0.89) \\
\bottomrule
\end{tabular}
\end{table}

% =====================================================================
\section{Generalization and honest scope}
% =====================================================================

The shape machinery is signal-agnostic: any response signal whose shape reflects the
underlying physical system, hydrogeology pumping tests, geothermal and thermal-response
tests, can be catalogued and conformally assigned by the same code. We exercise it on
real hydrogeology pumping-test field data (the Horkheim and Lauswiesen sites) alongside
the reservoir transients, all through one pipeline.

Scope is reported plainly. This is a reproduction plus extension of a published
pipeline, not a wholly new method for the base clustering and attribution, which are
credited to the reproduced work. The DFN pressure-transient simulation on the full
open-source flow engine is an offline dependency whose end-to-end integration is a
later phase; the results here rest on the paper's own committed corpus, analytic
dual-porosity families, and DFN ensembles, and the artifacts that lack a simulated
transient carry an explicit pending flag rather than a fabricated curve. The
dual-conformal coverage is a genuine trade: it buys physics-consistency and tighter
sets at the price of marginal coverage, and both sides are shown. Every number is a
committed, seed-deterministic artifact.

% =====================================================================
\section{Conclusion}
% =====================================================================

We independently reproduced, from permissively licensed components, a DFN-transient
GeoType pipeline for fractured-reservoir flow, recovering its clusters (silhouette up
to $0.86$) and its fracture-intensity attribution, and extended it with a
coverage-controlled Mondrian conformal assigner and, as the methodological
contribution, a dual-representation Mondrian conformal assigner that fuses shape and
physics-descriptor evidence to exclude right-shape-wrong-physics assignments. The
result is a reproducible, coverage-controlled way to place a new pressure transient in
a catalogue of flow behaviours and to say, with a guarantee, when it belongs to none.

% =====================================================================
\section*{Data and code availability}
% =====================================================================

Code, the committed derived artifacts, and the workbench are at
\url{https://github.com/fsantibanezleal/CAOS_RES_Pulso} (MIT); the reusable shape
machinery is \texttt{pygeotypes} (Apache-2.0). The reference corpus is the open 4TU
dataset of the reproduced study~\cite{data4tu} (GPL-3.0, kept in a local vault, never
committed); GPL workflow code is not vendored. Only compact derived artifacts are
committed.

% =====================================================================
\begin{thebibliography}{99}
\itemsep 0pt

\bibitem{kameltarghi2026} E.~Kamel Targhi, G.~Rongier, P.-O. Bruna, A.~Daniilidis,
and S.~Geiger, ``Unsupervised learning for geologically consistent fluid flow analysis
in fractured reservoirs,'' \emph{Comput. Geosci.}, vol.~30, 57, 2026,
doi:10.1007/s10596-026-10459-w.

\bibitem{warren1963} J.~E. Warren and P.~J. Root, ``The behavior of naturally
fractured reservoirs,'' \emph{SPE J.}, vol.~3, no.~3, pp.~245--255, 1963,
doi:10.2118/426-PA.

\bibitem{bourdet1989} D.~Bourdet, J.~A. Ayoub, and Y.~M. Pirard, ``Use of pressure
derivative in well-test interpretation,'' \emph{SPE Formation Eval.}, vol.~4, no.~2,
pp.~293--302, 1989, doi:10.2118/12777-PA.

\bibitem{stehfest1970} H.~Stehfest, ``Algorithm 368: Numerical inversion of Laplace
transforms,'' \emph{Commun. ACM}, vol.~13, no.~1, pp.~47--49, 1970,
doi:10.1145/361953.361969.

\bibitem{sakoe1978} H.~Sakoe and S.~Chiba, ``Dynamic programming algorithm
optimization for spoken word recognition,'' \emph{IEEE Trans. Acoust., Speech, Signal
Process.}, vol.~26, no.~1, pp.~43--49, 1978, doi:10.1109/TASSP.1978.1163055.

\bibitem{rousseeuw1987} P.~J. Rousseeuw, ``Silhouettes: A graphical aid to the
interpretation and validation of cluster analysis,'' \emph{J. Comput. Appl. Math.},
vol.~20, pp.~53--65, 1987, doi:10.1016/0377-0427(87)90125-7.

\bibitem{breiman2001} L.~Breiman, ``Random forests,'' \emph{Mach. Learn.}, vol.~45,
no.~1, pp.~5--32, 2001, doi:10.1023/A:1010933404324.

\bibitem{lundberg2020} S.~M. Lundberg \emph{et al.}, ``From local explanations to
global understanding with explainable AI for trees,'' \emph{Nat. Mach. Intell.},
vol.~2, pp.~56--67, 2020, doi:10.1038/s42256-019-0138-9.

\bibitem{vovk2005} V.~Vovk, A.~Gammerman, and G.~Shafer, \emph{Algorithmic Learning
in a Random World}. New York: Springer, 2005, doi:10.1007/b106715.

\bibitem{angelopoulos2023} A.~N. Angelopoulos and S.~Bates, ``Conformal prediction: A
gentle introduction,'' \emph{Found. Trends Mach. Learn.}, vol.~16, no.~4,
pp.~494--591, 2023, doi:10.1561/2200000101.

\bibitem{data4tu} E.~Kamel Targhi \emph{et al.}, ``GeoDFN networks and pressure
transients for geologically consistent fluid flow analysis (dataset),'' 4TU.ResearchData,
2026, doi:10.4121/8291d285-025d-4724-988d-fc747a578c0a.

\end{thebibliography}

\end{document}
